---
title: "04-EPID Integration"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = "../..")
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center',
                      message = FALSE, warning = FALSE, cache.lazy = TRUE,
                      cache = TRUE)
options(width = 1200)
```

## Introduction
In this script we aim to integrate the different datasets. According to the benchmarking paper by [Jinmiao Chen](https://genomebiology.biomedcentral.com/articles/10.1186/s13059-019-1850-9):

*Based on our results, Harmony, LIGER, and Seurat 3 are the recommended methods for batch integration. Due to its significantly shorter runtime,* ***Harmony*** *is recommended as the first method to try, with the other methods as viable alternatives.*

Therefore, we are going to use [Harmony](https://www.nature.com/articles/s41592-019-0619-0.epdf?shared_access_token=rDg_Rd07lrFXExt_ySj7V9RgN0jAjWel9jnR3ZoTv0NfDJkKCfDV_X9Mq3lweQmKiXEXxhrebQRjJEZdc-xNv6-7ZN1XotlD_mo5TSS4Z4eWn-kUo6mBwA5dEAKlTfR8OT6E10MZY_E-906ajbzvgg%3D%3D) to integrate the datasets.

[Harmony vignette](https://github.com/immunogenomics/harmony)
[Harmony Seurat wrapper](https://htmlpreview.github.io/?https://github.com/satijalab/seurat.wrappers/blob/master/docs/harmony.html)

## Libraries
```{r}
# library(devtools)
# install.packages("devtools")
# devtools::install_github("immunogenomics/harmony")

library(Seurat)
library(harmony)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(plotly)
```

## Set parameters
```{r}
source(here::here("misc/paths.R"))

dir.create(path = sprintf("%s/%s", clust_28, plt_dir),
           showWarnings = FALSE,
           recursive = TRUE)

dir.create(path = sprintf("%s/%s", clust_28, robj_dir),
           showWarnings = FALSE,
           recursive = TRUE)

# Set seed to make the script reproducible
set.seed(987)
```

## Load data
We are going to load the data processed and filtered in *03-QC_common.Rmd*.
With Harmony integration, we need to create only one Seurat object with all cells
```{r}
se_ls <- purrr::map(id_ls, ~ readRDS(file = here::here(glue::glue("{qc_28}/{robj_dir}/{.x}_qc.rds"))))

# merge all the objects in the list
merged_se <- purrr::reduce(se_ls, merge)
```

## Analysis

Add sample labels
```{r}
metadata <- merged_se@meta.data %>%
  tibble::rownames_to_column("barcode") %>%
  dplyr::left_join(id_df, c("orig.ident" = "gem_id")) %>%
  tibble::column_to_rownames("barcode")

merged_se@meta.data <- metadata
```

Preprocess data
```{r}
merged_se@assays$SCT <- NULL
Seurat::DefaultAssay(merged_se) <- "RNA"
merged_se <- Seurat::SCTransform(object = merged_se) %>%
  Seurat::RunPCA(., verbose = FALSE)
```

Lets see if the PCs merge well using the uncorrected PCs
```{r}
p1 <- Seurat::DimPlot(object = merged_se,
                      reduction = "pca",
                      pt.size = .1,
                      group.by = "sample_id")
p1

p2 <- Seurat::VlnPlot(object = merged_se,
              features = "PC_1",
              group.by = "sample_id",
              pt.size = .1)
p2
```

They all look quite good already prior to integration.
To make sure that just merging the data proves to be sufficient we want to visualize the data in a UMAP
```{r}
Seurat::ElbowPlot(merged_se, ndims = 50)
merged_test <- Seurat::FindNeighbors(merged_se, dims = 1:30) %>% 
  Seurat::FindClusters(., resolution = c(0.1))

merged_test <- Seurat::RunUMAP(merged_test, dims = 1:30)
```

```{r}
Seurat::DimPlot(object = merged_test,
                reduction = "umap",
                pt.size = .1,
                group.by = "orig.ident")
```

We see clusters but there is structure within the clusters corresponding to the data origin.
We will then proceed to integrate the data using Harmony

Start by selecting genes present in all the datasets
```{r}
# common_genes <- purrr::reduce(purrr::map(se_ls, rownames), intersect)
```

```{r}
merged_se <- harmony::RunHarmony(merged_se,
                                 group.by.vars = "orig.ident",
                                 assay.use = "SCT",
                                 reduction = "pca",
                                 dims = 1:30)
```

Remove previously individually-computed clusterings
```{r}
merged_se@meta.data <- merged_se@meta.data %>%
  dplyr::select(-starts_with("SCT_snn_res"))
```

Visualize integration result
```{r}
merged_se <- Seurat::FindNeighbors(merged_se,
                                   reduction = "harmony",
                                   dims = 1:30) %>%
            Seurat::FindClusters(resolution = c(0.01, 0.05, 0.1,
                                                0.25, 0.5, 0.75, 1))

merged_se <- Seurat::RunUMAP(merged_se,
                             reduction = "harmony",
                             dims = 1:30)

Seurat::Idents(merged_se) <- merged_se@meta.data[, "sample_id"]

int_plt <- Seurat::DimPlot(merged_se,
                           group.by = c("sample_id"),
                           pt.size = 0.1)
plotly::ggplotly(int_plt)
```

## Look at QC metrics
We are going to assess how the QC metrics look on the harmony embedding

### Basic QC
```{r fig.height=8, fig.width=8}
qc_plt <- Seurat::FeaturePlot(merged_se,
                              features = c("nCount_RNA",
                                           "nFeature_RNA",
                                           "pct_mt",
                                           "pct_ribo",
                                           "scrublet_doublet_scores"),
                              pt.size = .1)
qc_plt
```


```{r fig.height=10, fig.width=15}
dim_plt <- Seurat::DimPlot(merged_se,
                          group.by = c("SCT_snn_res.0.01",
                                       "SCT_snn_res.0.05",
                                       "SCT_snn_res.0.1",
                                       "SCT_snn_res.0.25",
                                       "SCT_snn_res.0.5",
                                       "SCT_snn_res.0.75",
                                       "SCT_snn_res.1"),
                          pt.size = .1)
dim_plt

Seurat::DimPlot(merged_se,
                group.by = c("SCT_snn_res.0.01",
                             "SCT_snn_res.0.25"),
                pt.size = 1,
                label = TRUE)

table(merged_se$SCT_snn_res.0.25)
```


```{r fig.height=8, fig.width=8}
plt1 <- Seurat::FeaturePlot(merged_se,
                            features = c("nCount_RNA"),
                            pt.size = 1)

plt2 <- Seurat::DimPlot(merged_se,
                        group.by = "Phase",
                        pt.size = 1)

plt3 <- Seurat::DimPlot(merged_se,
                        group.by = "SCT_snn_res.0.1",
                        pt.size = 1,
                        label = TRUE) + theme(legend.position = "none")

cowplot::plot_grid(plt1, plt2, plt3, ncol = 2)

# Seurat::FeaturePlot(merged_se, features = c("Vim", "Cd36"))
```

### Assess doublets
```{r}
scrublet_plt <- Seurat::FeaturePlot(merged_se,
                                    features = "scrublet_doublet_scores",
                                    pt.size = 1)
plotly::ggplotly(scrublet_plt)
```

## Save Harmony dataset
```{r}
# saveRDS(object = merged_se,
#         file = here::here(sprintf("%s/%s/harmony_se.rds",
#                        clust_28, robj_dir)))

"{clust_28}/{robj_dir}/harmony_se_epid20.rds" %>%
  glue::glue() %>%
  here::here() %>%
  saveRDS(
    object = merged_se,
    file = .)
```

## Session Info
```{r}
sessionInfo()
```
