---
author: "Marc Elosua-Bayes"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: html_document
params: 
    sample_id: "Default!"
title: "`r sprintf('%s Clustering', params$sample_id)`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Introduction
This script is set to carry out data scaling and normalization and clustering at different resolutions.
In this script we use the data coming from *02-QC/02-QC_common.Rmd*.

## Libraries

```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{clust_27}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(
    path = .,
    showWarnings = FALSE,
    recursive = TRUE)

"{clust_27}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(
    path = .,
    showWarnings = FALSE,
    recursive = TRUE)
```

```{r}
# sample_id <- params$sample_id
sample_id <- "s2yq8o_t5hk8u"
donor_id <- qc_dict[[sample_id]][["sample_id"]]
```

## Load data
We have 4 different datasets that we are going to analyze separately. In each Rmarkdown notebook we analyze 1 dataset.
```{r}
se_obj <- "{qc_27}/{robj_dir}/qc_se_{sample_id}.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)
```

## Analysis

### Scale and Normalize
```{r}
se_obj <- se_obj %>% 
  Seurat::NormalizeData(., verbose = FALSE) %>%
  Seurat::FindVariableFeatures(., verbose = FALSE) %>%
  Seurat::ScaleData(., verbose = FALSE) %>%
  Seurat::RunPCA(., verbose = FALSE)
```

### Clustering + non-linear dimensionality embedding
```{r}
Seurat::ElbowPlot(se_obj)

se_obj <- Seurat::FindNeighbors(se_obj,
                                   reduction = "pca",
                                   dims = 1:10,
                                   verbose = FALSE) %>% 
  Seurat::FindClusters(resolution = c(0.1, 0.3, 0.5, 0.75, 1))

se_obj <- Seurat::RunUMAP(se_obj,
                          reduction = "pca",
                          dims = 1:10,
                          verbose = FALSE)
```

Visualize the clustering
```{r fig.height=8, fig.width=10}
Seurat::DimPlot(object = se_obj,
                group.by = c("Spatial_snn_res.0.1",
                             "Spatial_snn_res.0.3",
                             "Spatial_snn_res.0.5",
                             "Spatial_snn_res.0.75",
                             "Spatial_snn_res.1"),
                label = TRUE)
```

```{r fig.height=10, fig.width=14}
Seurat::SpatialDimPlot(object = se_obj,
                group.by = c("Spatial_snn_res.0.1",
                             "Spatial_snn_res.0.3",
                             "Spatial_snn_res.0.5",
                             "Spatial_snn_res.0.75",
                             "Spatial_snn_res.1"))

```

## Save object
```{r}
"{clust_27}/{robj_dir}/clust_se_{sample_id}.rds" %>%
  glue::glue() %>%
  here::here() %>%
  saveRDS(object = se_obj, file = .)
```

We are also going to save the files to be used in the shiny app. We can find the *seurat2shiny* function [here](https://github.com/Single-Cell-Genomics-Group-CNAG-CRG/shiny-annotation/blob/main/seurat2shiny.R)
```{r eval=FALSE}
seurat2shiny = function(
    object                         ,
    tech      = c("sc", "sp")      ,
    assay     = object@active.assay,
    slot      = "data"             ,
    reduction = "umap"             ,
    image     = NULL               ,
    asfactors = NULL               ,
    save      = TRUE               ,
    path      = "."                  # path = getwd()
) {
    suppressMessages( library(Seurat) );

    # Input check.
    if ( ! is(object, "Seurat") )
        stop("'object' is not a Seurat object.");

    if ( ! assay %in% Seurat::Assays(object) )
        stop("'assay' not in the Seurat object's available assays.");

    if ( tech == "sc" & ! (reduction %in% names(object@reductions)) )
        stop("'reduction' not in the Seurat object's available reductions.");

    if ( ! slot %in% c("counts", "data", "scale.data") )
        stop("'slot' not in the Seurat object's available slots.");
    
    if ( ! tech %in% c("sc", "sp") )
        stop("tech must be sc or sp.");
    
    
    # Check Which technology it is processing
    if (tech == "sc") {
        # Extract 2D coordinates.
        embeds <- as.data.frame(object@reductions[[reduction]]@cell.embeddings);
        names(embeds) <- c("coord_x", "coord_y");
    } else if (tech == "sp") {
        # If the iimage is null select the first one
        if (is.null(image)) {
            image <- names(object@images)[1]
            warning(sprintf("image is not set, we will use %s", image))
        } 
        
        embeds <- data.frame(object@images[[image]]@coordinates[, c("imagerow", "imagecol")])
        colnames(embeds) <- c("coord_y", "coord_x");
        
        # Inverse coord_y
        embeds$coord_y <- - embeds$coord_y
    }
    

    # Join metadata with coordinates.
    metadata <- object@meta.data;

    for (col in asfactors) {
        metadata[[col]] <- as.factor(metadata[[col]]);
    };

    metadata <- merge(x = metadata, y = embeds, by = "row.names");
    names(metadata)[1] <-  "barcode"; # names(metadata)[names(metadata) == "Row.names"] = "barcode";
    rownames(metadata) <- metadata$barcode

    # Extract expression data.
    # expression = as.matrix( Seurat::GetAssayData(object = object, slot = slot, assay = assay) );
    expression = Seurat::GetAssayData(object = object, slot = slot, assay = assay);

    if ( ! identical( as.character(metadata$barcode), colnames(expression) ) )
        warning("Cells in metadata and expression matrix do not match.");

    if (save) {
        saveRDS( object = metadata  , file = paste0(path, "/metadata.rds"  ) );
        saveRDS( object = expression, file = paste0(path, "/expression.rds") );
    };

    invisible(
        list(metadata = metadata, expression = expression)
    );
}
```

```{r eval=FALSE}
list_shiny <- seurat2shiny(object = se_obj,
                           tech = "sp",
                           image = sample_id,
                           assay = "Spatial",
                           slot = "data")

saveRDS(object = list_shiny[[1]],
        file = here::here(sprintf("%s/%s/metadata_%s.RDS",
                                  clust_27, robj_dir, donor_id)))
saveRDS(object = list_shiny[[2]],
        file = here::here(sprintf("%s/%s/expression_%s.RDS",
                                  clust_27, robj_dir, donor_id)))
```

## Session Info
```{r}
sessionInfo()
```
