---
author: "Marc Elosua-Bayes"
title: "Extended Figure 25-A,B,C"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Introduction
In this Rmarkdown we are going to plot panels D, G & H.
In this script we will use the the scRNAseq data from *sc_analysis/04-annotation/07-join_annotation.Rmd*.

## Libraries

```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(glue)
library(stringr)
library(readr)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{fig_pt}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(
    path = .,
    showWarnings = FALSE,
    recursive = TRUE)

"{fig_pt}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(
    path = .,
    showWarnings = FALSE,
    recursive = TRUE)

img_order <- c("uzntl1_fkclxp", "k886qc_rqix54","s2yq8o_t5hk8u", "ivuznh_s8xyfv")

SpatialColors <- colorRampPalette(colors = rev(x = brewer.pal(n = 11, name = "Spectral")))
```

## Load data
Load Visium data
```{r}
# 07-sc_mapping_viz.Rmd
# se_obj <- "{map_27}/{robj_dir}/se_deconv_{sample_id}_epid20_pre-rotation.rds"

sp_ls <- lapply(id_sp_df$gem_id, function(id) {

  se_obj <- "{map_27}/{robj_dir}/se_deconv_{id}_epid20.rds" %>%
    glue::glue() %>%
    here::here() %>%
    readRDS(file = .)

  return(se_obj)
})

se_obj <- merge(sp_ls[[1]], y = sp_ls[2:length(sp_ls)],
                add.cell.ids = id_sp_df$gem_id,
                project = "Gloria-Salva")
```

## Panels

### Panel A
In this panel we show the HE images and the mouse percentage per spot.

```{r fig.width=16, fig.height=4}
row1_a <- Seurat::SpatialPlot(
  object = se_obj,
  features = "GRCh38-AGL",
  alpha = c(0, 0),
  images = img_order,
  crop = FALSE,
  image.alpha = 1,
  pt.size.factor = 1.25) &
  Seurat::NoLegend() &
  ggplot2::labs(title = "")

row1_a

row2_a <- Seurat::SpatialPlot(
  object = se_obj,
  features = c("percent.mouse"),
  images = img_order,
  crop = FALSE,
  image.alpha = 0,
  pt.size.factor = 1.25) &
  ggplot2::scale_fill_gradientn(
    colours = SpatialColors(n = 100),
    limits = c(0, 1)) &
  ggplot2::labs(title = "")

row2_a <- ggpubr::ggarrange(row2_a[[1]], row2_a[[2]], row2_a[[3]], row2_a[[4]],
                              ncol = 4, common.legend = TRUE, legend = "right")
row2_a

"{fig_pt}/{plt_dir}/Extended_25-A1.svg" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = row1_a,
    base_height = 4,
    base_width = 16)

"{fig_pt}/{plt_dir}/Extended_25-A2.svg" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = row2_a,
    base_height = 4,
    base_width = 16)
```

### Panel B
In this image we show the tumor-associated Schwann cells predicted proportion along with the mouse percentage and the tissue stratification. 

Stratify the tissue
```{r}
se_obj@meta.data <- se_obj@meta.data %>%
  dplyr::mutate(
    stratification = dplyr::case_when(
      sample_id == "GP027_PKLO_Palm_Diet" & Spatial_snn_res.0.5 %in% c(0, 2, 5) ~ "Tumour",
      sample_id == "GP027_PKLO_Palm_Diet" & Spatial_snn_res.0.5 == 3 ~ "Tumour Front",
      sample_id == "GP027_PKLO_Palm_Diet" & Spatial_snn_res.0.5 %in% c(1, 4, 6, 7, 8) ~ "Healthy",
      sample_id == "GP20004_dKDCD36_CT_Diet" & Spatial_snn_res.0.1 == 1 ~ "Tumour",
      sample_id == "GP20004_dKDCD36_CT_Diet" & Spatial_snn_res.0.1 == 2 ~ "Tumour Front",
      sample_id == "GP20004_dKDCD36_CT_Diet" & Spatial_snn_res.0.1 %in% c(0, 3) ~ "Healthy",
      sample_id == "GP029_PLKO_CT_Diet" & Spatial_snn_res.0.3 == 4 ~ "Tumour",
      sample_id == "GP029_PLKO_CT_Diet" & Spatial_snn_res.0.3 == 3 ~ "Tumour Front",
      sample_id == "GP029_PLKO_CT_Diet" & ! Spatial_snn_res.0.3 %in% c(3, 4) ~ "Healthy",
      sample_id == "GP024_dKDCD36_Palm_Diet" & Spatial_snn_res.1 %in% c(3, 7) ~ "Tumour",
      sample_id == "GP024_dKDCD36_Palm_Diet" & Spatial_snn_res.1 %in% c(6, 8) ~ "Tumour Front",
      sample_id == "GP024_dKDCD36_Palm_Diet" & ! Spatial_snn_res.1 %in% c(3, 6, 7, 8) ~ "Healthy"
      ),
    stratification = factor(stratification,
                            levels = c("Healthy", "Tumour Front", "Tumour"))
    )
```

Plot arrangement
```{r}
b_ls <- lapply(img_order, function(img) {
  
  # Schwann cells
  tmp1 <- Seurat::SpatialPlot(
    object = se_obj,
    features = c("Tumour-associated Schwann Cells"),
    images = img,
    crop = FALSE,
    image.alpha = 0,
    pt.size.factor = 1.25) &
    ggplot2::scale_fill_gradientn(
      colours = SpatialColors(n = 100),
      limits = c(0, max(se_obj$`Tumour-associated Schwann Cells`)))
  
  # percent.mouse
  tmp2 <- Seurat::SpatialPlot(
    object = se_obj,
    features = c("percent.mouse"),
    images = img,
    crop = FALSE,
    image.alpha = 0,
    pt.size.factor = 1.25) &
    ggplot2::scale_fill_gradientn(
      colours = SpatialColors(n = 100),
      limits = c(0, 1))
  
  # percent.mouse
  tmp3 <- Seurat::SpatialPlot(
    object = se_obj,
    group.by = "stratification",
    images = img,
    crop = FALSE,
    image.alpha = 0,
    pt.size.factor = 1.25) +
    ggplot2::scale_fill_manual(
      values = c("#009E73", "#E69F00", "#D55E00"),
      breaks = c("Healthy", "Tumour Front", "Tumour")) +
    ggplot2::theme(legend.position = "top") +
    ggplot2::guides(
      fill = ggplot2::guide_legend(override.aes = list(size = 5)))
  
  cowplot::plot_grid(
    plotlist = list(tmp1, tmp2, tmp3),
    align = "hv",
    axis = "trbl",
    nrow = 1)

})
```

Join the sub-panels of panel B
```{r fig.width=30, fig.height=12}
panel_b <- plot_grid(plotlist = b_ls, ncol = 2, align = "hv", axis = "trbl")
panel_b

"{fig_pt}/{plt_dir}/Extended_25-B.svg" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = panel_b,
    base_height = 12,
    base_width = 30)
```

### Panel C
In this panel we show the violin plots of the predicted proportions of tumour-associated Schwann cells stratified by healthy, tumour front and tumour.

```{r fig.width=12, fig.height=9}
my_comparisons <- list(
    c("Tumour", "Tumour Front"),
    c("Tumour", "Healthy"),
    c("Tumour Front", "Healthy")
    )

panel_c <- se_obj@meta.data %>%
  dplyr::mutate(
    sample_id = factor(sample_id,
                       levels = c( "GP027_PKLO_Palm_Diet", "GP029_PLKO_CT_Diet",
                                   "GP024_dKDCD36_Palm_Diet", "GP20004_dKDCD36_CT_Diet"))
  ) %>%
  ggplot2::ggplot(.,
  ggplot2::aes(x = stratification, y = `Tumour-associated Schwann Cells`)) +
  ggplot2::geom_violin(
    alpha = 0.7,
    ggplot2::aes(fill = stratification, color = stratification)) +
  ggplot2::geom_jitter(ggplot2::aes(color = stratification)) +
  ggplot2::facet_wrap(.~sample_id, scales = "free") +
  ggplot2::labs(
    x = "Tissue Stratification",
    y = "Proportion of Tumour-associated Schwann Cells") +
  ggplot2::theme_classic() +
  ggpubr::stat_compare_means(
    label = "p.format",
    comparisons = my_comparisons,
    p.adjust.method = "bonferroni",
    ) +
  # ggpubr::stat_compare_means(label.y = c(0.75, 0.5, 0.85, 0.75)) + # Add global p-value
  ggplot2::scale_fill_manual(values = c("#009E73", "#E69F00", "#D55E00")) +
  ggplot2::scale_color_manual(values = c("#009E73", "#E69F00", "#D55E00")) +
  ggplot2::theme(legend.title = element_blank())

panel_c

"{fig_pt}/{plt_dir}/Extended_25-C.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = panel_c,
    base_height = 9,
    base_width = 12)
```

## Session Info
```{r}
sessionInfo()
```
