---
author: "Marc Elosua-Bayes"
title: "Extended Figure 23-A-C"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Introduction
In this Rmarkdown we are going to plot panels D, G & H.
In this script we will use the the scRNAseq data from *sc_analysis/04-annotation/07-join_annotation.Rmd*.

## Libraries

```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(glue)
library(stringr)
library(readr)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{fig_pt}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(
    path = .,
    showWarnings = FALSE,
    recursive = TRUE)

"{fig_pt}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(
    path = .,
    showWarnings = FALSE,
    recursive = TRUE)
```

## Load data
Load scRNAseq data
```{r}
# *sc_analysis/04-annotation/07-join_annotation.Rmd
sc_obj <- "{anot_28}/{robj_dir}/harmony_se_annot.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)

mask_remove <- ! sc_obj$specific_annot %in% c("Proliferating Basal Cells",
                                              "Basal Cells", "Erithrocytes")

```

## Panels

### Panel A
In this panel we show the UMAP with the genes of interest identifying Glial and Glial progenitors.

```{r fig.width=30, fig.height=10}
feats_a <- c("Slc1a3", "Scn7a", "Slc16a2", "Dio2", "Slc6a2",
           "Plp1", "Aldh1l1", "Pdgfra", "Cspg4")

panel_a <- Seurat::FeaturePlot(
  object = sc_obj[, mask_remove],
  features = c(feats_a),
  ncol = 5,
  pt.size = 0.1) &
  ggplot2::scale_color_gradient(low = "#FFFF80FF",
                              high = "#FF0000FF")

panel_a

"{fig_pt}/{plt_dir}/Extended_23-A.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = panel_a,
    base_height = 10,
    base_width = 30)
```

## Panel B
In this panel we show the UMAP with the genes of interest identifying Perineuronal Net - ECM Constituent.

```{r fig.width=25, fig.height=10}
feats_b <- c("Bgn", "Tnn", "Lum", "Ngf", "Sulf1", "Mmp2", "Timp1", "Has2")

panel_b <- Seurat::FeaturePlot(
  object = sc_obj[, mask_remove],
  features = c(feats_b),
  ncol = 4,
  pt.size = 0.1) &
  ggplot2::scale_color_gradient(low = "#FFFF80FF",
                              high = "#FF0000FF")

panel_b

"{fig_pt}/{plt_dir}/Extended_23-B.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = panel_b,
    base_height = 10,
    base_width = 25)
```

## Panel C
In this panel we show the UMAP with the genes of interest identifying Perineuronal Net - ECM Constituent.

```{r fig.width=25, fig.height=10}
feats_c <- c("Enpp2", "Stmn4", "Saa3", "Il33", "Lpar1", "Gpm6b", "Olfml3", "Ccl11")

panel_c <- Seurat::FeaturePlot(
  object = sc_obj[, mask_remove],
  features = c(feats_c),
  ncol = 4,
  pt.size = 0.1) &
  ggplot2::scale_color_gradient(low = "#FFFF80FF",
                              high = "#FF0000FF")

panel_c

"{fig_pt}/{plt_dir}/Extended_23-C.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = panel_c,
    base_height = 10,
    base_width = 25)
```

## Session Info
```{r}
sessionInfo()
```
