---
author: "Marc Elosua-Bayes"
title: "Extended Figure 26"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Introduction
In this Rmarkdown we are going to plot panels A, B & C.
In this script we will use the Visium data coming from *spatial_analysis/05-sc_mapping/07-sc_mapping_viz.Rmd*.

## Libraries

```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(glue)
library(stringr)
library(readr)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{fig_pt}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(
    path = .,
    showWarnings = FALSE,
    recursive = TRUE)

"{fig_pt}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(
    path = .,
    showWarnings = FALSE,
    recursive = TRUE)

img_order <- c("uzntl1_fkclxp", "k886qc_rqix54","s2yq8o_t5hk8u", "ivuznh_s8xyfv")

SpatialColors <- colorRampPalette(colors = rev(x = brewer.pal(n = 11, name = "Spectral")))
```

## Load data
Load Visium data
```{r}
# 07-sc_mapping_viz.Rmd
# se_obj <- "{map_27}/{robj_dir}/se_deconv_{sample_id}_epid20_pre-rotation.rds"

sp_ls <- lapply(id_sp_df$gem_id, function(id) {

  se_obj <- "{map_27}/{robj_dir}/se_deconv_{id}_epid20.rds" %>%
    glue::glue() %>%
    here::here() %>%
    readRDS(file = .)

  return(se_obj)
})

se_obj <- merge(sp_ls[[1]], y = sp_ls[2:length(sp_ls)],
                add.cell.ids = id_sp_df$gem_id,
                project = "Gloria-Salva")
```

## Panels

### Panel A
In this panel we will show marker genes related to Glial and Glial progenitor on the Visium slide.

```{r}
feats_a <- c("mm10---S100b", "mm10---Scn7a", "mm10---Sox10")

a_ls <- lapply(feats_a, function(feat) {
  
  tmp_row <- lapply(img_order, function(img) {
    
    tmp <- Seurat::SpatialPlot(
      object = se_obj,
      features = feat,
      images = img,
      image.alpha = 0,
      crop = FALSE,
      pt.size.factor = 1.25) &
      ggplot2::scale_fill_gradientn(
        colours = SpatialColors(n = 100),
        limits = c(0, max(se_obj@assays$Spatial@data[feat, ])))
    
  }) %>%
    ggpubr::ggarrange(
      plotlist = .,
      nrow = 1,
      common.legend = TRUE,
      legend = "right")
  })
```

Join panel A
```{r}
panel_a <- cowplot::plot_grid(
  plotlist = a_ls,
  align = "hv",
  axis = "trbl",
  nrow = 3)

panel_a

"{fig_pt}/{plt_dir}/Extended_26-A.svg" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = panel_a,
    base_height = 10,
    base_width = 15)
```

### Panel B
In this panel we will show marker genes related to Perineuronal Net - ECM constituent on the Visium slide.

```{r}
feats_b <- c("mm10---Sulf1", "mm10---Has1", "mm10---Timp1", "mm10---Cspg4")

b_ls <- lapply(feats_b, function(feat) {
  
  tmp_row <- lapply(img_order, function(img) {
    
    tmp <- Seurat::SpatialPlot(
      object = se_obj,
      features = feat,
      images = img,
      image.alpha = 0,
      crop = FALSE,
      pt.size.factor = 1.25) &
      ggplot2::scale_fill_gradientn(
        colours = SpatialColors(n = 100),
        limits = c(0, max(se_obj@assays$Spatial@data[feat, ])))
    
  }) %>%
    ggpubr::ggarrange(
      plotlist = .,
      nrow = 1,
      common.legend = TRUE,
      legend = "right")
  })

```

Join panel B
```{r}
panel_b <- cowplot::plot_grid(
  plotlist = b_ls,
  align = "hv",
  axis = "trbl",
  nrow = 4)

panel_b

"{fig_pt}/{plt_dir}/Extended_26-B.svg" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = panel_b,
    base_height = 15,
    base_width = 18)
```

### Panel C
In this panel we will show marker genes related to Nerve injury and Pro-regenerative state on the Visium slide.

```{r}
feats_c <- c("mm10---Il33", "mm10---Ccl11", "mm10---Lpar1", "mm10---Enpp2")

c_ls <- lapply(feats_c, function(feat) {
  
  tmp_row <- lapply(img_order, function(img) {
    
    tmp <- Seurat::SpatialPlot(
      object = se_obj,
      features = feat,
      images = img,
      image.alpha = 0,
      crop = FALSE,
      pt.size.factor = 1.25) &
      ggplot2::scale_fill_gradientn(
        colours = SpatialColors(n = 100),
        limits = c(0, max(se_obj@assays$Spatial@data[feat, ])))
    
  }) %>%
    ggpubr::ggarrange(
      plotlist = .,
      nrow = 1,
      common.legend = TRUE,
      legend = "right")
  })

```

Join panel C
```{r}
panel_c <- cowplot::plot_grid(
  plotlist = c_ls,
  align = "hv",
  axis = "trbl",
  nrow = 4)

panel_c

"{fig_pt}/{plt_dir}/Extended_26-C.svg" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = panel_c,
    base_height = 15,
    base_width = 18)
```

## Session Info
```{r}
sessionInfo()
```
