---
author: "Marc Elosua-Bayes"
title: "Main Figure 3-D,G,H"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Introduction
In this Rmarkdown we are going to plot panels D, G & H.
In this script we will use the Visium data coming from *spatial_analysis/05-sc_mapping/07-sc_mapping_viz.Rmd* and the scRNAseq data from *sc_analysis/04-annotation/07-join_annotation.Rmd*.

## Libraries

```{r}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(glue)
library(stringr)
library(readr)
```

## Setting parameters
Loading necessary paths and parameters
```{r}
set.seed(123)
source(here::here("misc/paths.R"))
source(here::here("utils/bin.R"))

"{fig_pt}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(
    path = .,
    showWarnings = FALSE,
    recursive = TRUE)

"{fig_pt}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(
    path = .,
    showWarnings = FALSE,
    recursive = TRUE)

SpatialColors <- colorRampPalette(colors = rev(x = brewer.pal(n = 11, name = "Spectral")))
```

## Load data
Load Visum and scRNAseq data

```{r}
# 07-sc_mapping_viz.Rmd
# se_obj <- "{map_27}/{robj_dir}/se_deconv_{sample_id}_epid20_pre-rotation.rds"

sp_ls <- lapply(id_sp_df$gem_id, function(id) {

  se_obj <- "{map_27}/{robj_dir}/se_deconv_{id}_epid20.rds" %>%
    glue::glue() %>%
    here::here() %>%
    readRDS(file = .)

  return(se_obj)
})

se_obj <- merge(sp_ls[[1]], y = sp_ls[2:length(sp_ls)],
                add.cell.ids = id_sp_df$gem_id,
                project = "Gloria-Salva")

# *sc_analysis/04-annotation/07-join_annotation.Rmd
sc_obj <- "{anot_28}/{robj_dir}/harmony_se_annot.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)
```

## Panels

### Panel D
In this panel we show the UMAP with the all cell identity populations

```{r fig.width=12, fig.height=8}
panel_d <- Seurat::DimPlot(
  object = sc_obj,
  group.by = c("specific_annot")) +
  ggplot2::labs(title = "")

panel_d

"{fig_pt}/{plt_dir}/Main_Figure3-D.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = panel_d,
    base_height = 8,
    base_width = 12)
```

### Panel F
In this panel we show genes of interest on the cell identity's of interest.
```{r fig.width=12, fig.height=8}
mask_remove <- ! sc_obj$specific_annot %in% c("Proliferating Basal Cells",
                                              "Basal Cells", "Erithrocytes")
panel_f <- Seurat::FeaturePlot(
  object = sc_obj[, mask_remove],
  features =  c(
    "Apod", # Nerve regeneration
    "Nrep", # Nerve regeneration
    "Ncam1", # Non-myelinating-Immature Schwann cells
    "Vcan", # Extracellular matrix / Perineuronal net constituent
    "Has1", # Extracellular matrix / Perineuronal net constituent
    "Tnc" # Extracellular matrix / Perineuronal net constituent
    ),
  ncol = 3, pt.size = 0.25) &
  ggplot2::scale_color_gradient(
    low = "yellow",
    high = "red")

panel_f

"{fig_pt}/{plt_dir}/Main_Figure3-F.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = panel_f,
    base_height = 8,
    base_width = 12)
```

### Panel G
In this panel we want to show the mouse % along with the predicted proportion of tumour-associated Schwann cells in the Visium slides.

```{r fig.width=16, fig.height=4}
row1 <- Seurat::SpatialPlot(
  object = se_obj,
  features = c("Tumour-associated Schwann Cells"),
  images = c("uzntl1_fkclxp", "k886qc_rqix54","s2yq8o_t5hk8u", "ivuznh_s8xyfv"),
  crop = FALSE,
  pt.size.factor = 1.25,
  image.alpha = 0) &
  ggplot2::scale_fill_gradientn(
    colours = SpatialColors(n = 100),
    limits = c(0, max(se_obj$`Tumour-associated Schwann Cells`)))

row1_mod <- ggpubr::ggarrange(row1[[1]], row1[[2]], row1[[3]], row1[[4]],
                              ncol = 4, common.legend = TRUE, legend = "right")
row1_mod

row2 <- Seurat::SpatialPlot(
  object = se_obj,
  features = c("percent.mouse"),
  images = c("uzntl1_fkclxp", "k886qc_rqix54","s2yq8o_t5hk8u", "ivuznh_s8xyfv"),
  crop = FALSE,
  pt.size.factor = 1.25,
  image.alpha = 0) &
  ggplot2::scale_fill_gradientn(
    colours = SpatialColors(n = 100),
    limits = c(0, 1))

row2_mod <- ggpubr::ggarrange(row2[[1]], row2[[2]], row2[[3]], row2[[4]],
                              ncol = 4, common.legend = TRUE, legend = "right")
row2_mod

"{fig_pt}/{plt_dir}/Main_Figure3-G1.svg" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = row1_mod,
    base_height = 4,
    base_width = 16)


"{fig_pt}/{plt_dir}/Main_Figure3-G2.svg" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = row2_mod,
    base_height = 4,
    base_width = 16)

```

### Panel H
In this panel we want to show genes of interest in the spatial slides
```{r fig.width=16, fig.height=12}
plt_ls <- lapply(c("GRCh38-CD36", "mm10---Vcan", "mm10---Tnn"), function(feat) {
  
  tmp <- Seurat::SpatialPlot(
  object = se_obj,
  features = feat,
  images = c("uzntl1_fkclxp", "k886qc_rqix54","s2yq8o_t5hk8u", "ivuznh_s8xyfv"),
  crop = FALSE,
  pt.size.factor = 1.25,
  image.alpha = 0) &
  ggplot2::scale_fill_gradientn(
    colours = SpatialColors(n = 100),
    limits = c(0, max(se_obj@assays$Spatial@data[feat, ])))

  tmp <- ggpubr::ggarrange(
    tmp[[1]], tmp[[2]], tmp[[3]], tmp[[4]],
    ncol = 4, common.legend = TRUE, legend = "right")

})

panel_h <- cowplot::plot_grid(
  plotlist = plt_ls,
  align = "hv",
  axis = "trbl",
  nrow = 3)

panel_h

"{fig_pt}/{plt_dir}/Main_Figure3-H.svg" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = panel_h,
    base_height = 12,
    base_width = 16)
```

## Session Info
```{r}
sessionInfo()
```

