---
title: "Palm diet comparison"
author: "Marc Elosua"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# knitr::opts_knit$set(root.dir = "../..")
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center', 
                      message = FALSE, warning = FALSE)
options(width = 1200)
```

## Introduction
In this Rmarkdown document we are going to compare the cell type proportions and DE genes between experimental conditions.

## Libraries
```{r warning = FALSE, message = FALSE}
library(dplyr)
library(Seurat)
library(DT)
library(ggplot2)
# library(ggVennDiagram)
library(emmeans)
library(ggrepel)
```

## Parameters
Here we will define and load objects and functions that will be used throughout the document.
```{r}
set.seed(321)
source(here::here("utils/bin.R"))
source(here::here("misc/paths.R"))

"{palm_28}/{plt_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(
    path = .,
    showWarnings = FALSE,
    recursive = TRUE)

"{palm_28}/{robj_dir}" %>%
  glue::glue() %>%
  here::here() %>%
  dir.create(
    path = .,
    showWarnings = FALSE,
    recursive = TRUE)
```

## Load data
The data used in this Rmarkdown document comes from...
```{r}
# annot_se <- "{anot_28}/{robj_dir}/annotated_se.rds" %>%
annot_se <- "{anot_28}/{robj_dir}/harmony_se_annot.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)
```

## Analysis

We start by looking at the cell-composition barplot between conditions to assess if there are different cell type proportions.
```{r fig.height = 8, fig.width = 15}
colourCount <- length(unique(annot_se$specific_annot))
getPalette <- colorRampPalette(RColorBrewer::brewer.pal(9, "Set3"))

annot_se@meta.data %>%
  dplyr::group_by(sample_id) %>%
  dplyr::mutate(n_group = n()) %>%
  dplyr::group_by(sample_id, specific_annot ,n_group) %>%
  dplyr::summarise(n = n()) %>%
  dplyr::mutate(x_tick = stringr::str_c(sample_id, n_group, sep = "\n")) %>% 
  ggplot2::ggplot(ggplot2::aes(x = x_tick,
             y = n,
             fill = specific_annot
             )) + 
    ggplot2::geom_bar(position = "fill",
             stat = "identity") +
  ggpubr::theme_pubr() +
  ggplot2::labs(
    x = "Diet",
    y = "Proportion") +
  ggplot2::scale_fill_manual(values = getPalette(colourCount)) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 60,
                                   vjust = 1,
                                   hjust = 1),
        legend.title = ggplot2::element_blank())
```

Statistically assess the difference between cell type proportions
We follow the approached presented by Valentine Svensson [here](https://www.nxn.se/valent/2020/11/28/s9jjv32ogiplagwx8xrkjk532p7k28).
There a binomial linear model is used to analyse counts of repeated observations of binary choices.
```{r}
# Make sure all the cell types are there for all experimental conditions, add 0 to those missing
all_ct <- unique(annot_se@meta.data$specific_annot)
all_sample <- unique(annot_se@meta.data$sample_id)
df <- data.frame("sample_id" = sort(rep(x = all_sample, length(all_ct))),
                 "specific_annot" = rep(x = all_ct, length(all_sample)),
                 "count_null" = 0,
                 stringsAsFactors = FALSE)

# df <- df %>%
#   # dplyr::left_join(annot_se@meta.data, ) %>%
#   dplyr::mutate(
#     count = dplyr::if_else(is.na(count), 0L , count),
#     total = total,
#     other = total - count
#   ) %>% data.frame()


metadata <- annot_se@meta.data %>%
  dplyr::right_join(df, by = c("specific_annot", "sample_id")) %>% 
  dplyr::select(sample_id, specific_annot) %>%
  # Get how many cells per cell type for each sample
  dplyr::group_by(sample_id, specific_annot) %>%
  dplyr::summarise(count = n()) %>%
  dplyr::ungroup() %>%
  # Get total number of cells for each sample
  dplyr::group_by(sample_id) %>%
  dplyr::mutate(
    # 
    count = dplyr::if_else(count == 1, 0L, count),
    total = sum(count),
    other = total - count,
    experiment = dplyr::case_when(
      stringr::str_detect(string = sample_id,
                          pattern = stringr::regex("plko", ignore_case = TRUE)) ~ "PLKO",
      stringr::str_detect(string = sample_id,
                            pattern = stringr::regex("shcd36", ignore_case = TRUE)) ~ "shcD36",
      stringr::str_detect(string = sample_id,
                          pattern = stringr::regex("shgal", ignore_case = TRUE)) ~ "shGal",
      stringr::str_detect(string = sample_id,
                          pattern = stringr::regex("Normal", ignore_case = TRUE)) ~ "Normal",
      TRUE ~ NA_character_),
    diet = dplyr::case_when(
      stringr::str_detect(string = sample_id,
                          pattern = stringr::regex("ct|control", ignore_case = TRUE)) ~ "Control",
      stringr::str_detect(string = sample_id,
                            pattern = stringr::regex("palm", ignore_case = TRUE)) ~ "Palm",
      TRUE ~ "Control"),
    facs = dplyr::if_else(sample_id %in% c("25_Plko_ct_diet",
                                           "25_Plko_Palmdiet",
                                           "25_shcD36_ctdiet",
                                           "25_shcD36_PALM_diet"),
                          "No-FACS", "EpCAM-FACS"),
    pairs = glue::glue("{experiment}-{facs}")
    ) %>%
  dplyr::ungroup()

# metadata[metadata$sample_id == "25_Plko_ct_diet", "specific_annot"]
# metadata[metadata$sample_id == "Normaltongue", "specific_annot"]
```

To compute the differences between the number of cell types in 2 conditions we are going to use the *Chi-square test*. This test examines whether rows and columns of a contingency table are statistically significantly associated. \
* **Null hypothesis (H0)**: the row and the column variables of the contingency table are independent. \
* **Alternative hypothesis (H1)**: row and column variables are dependent. \

1st - For each cell of the table, we have to calculate the expected value under null hypothesis. \
For a given cell, the expected value is calculated as follow: \
\begin{align}
e=\frac{row.sum*col.sum}{grand.total}
\end{align}
Then the Chi-square statistic is computed as:
\begin{align}
X^2=\sum\frac{(o - e)^2}{e}
\end{align}

```{r}
options(pillar.sigfig = 3)

pairs_ls <- metadata %>%
  dplyr::filter(pairs != "Normal-EpCAM-FACS") %>%
  dplyr::pull(pairs) %>%
  unique()

# Iterate over the pairs of experimemts
chisq_ls <- lapply(pairs_ls, function(pair) {
  print(pair)
  # Extract sample id for those conditions
  sample_id_vec <- metadata %>%
    dplyr::filter(pairs == pair) %>%
    dplyr::pull(sample_id) %>%
    unique()
  
  # Subet both cell-type compositions
  diet_1 <- metadata %>%
  dplyr::filter(sample_id == sample_id_vec[1]) %>%
  dplyr::select(specific_annot, count) %>%
  dplyr::mutate(count_ct_diet = count) %>%
  dplyr::select(-count)
  
  diet_2 <- metadata %>%
    dplyr::filter(sample_id == sample_id_vec[2]) %>%
    dplyr::select(specific_annot, count) %>%
    dplyr::mutate(count_palm_diet = count) %>%
    dplyr::select(-count)
  
  # join cell type numbers
  diet_cell_types <- dplyr::left_join(
    diet_1,
    diet_2,
    by = "specific_annot") %>%
  tibble::column_to_rownames("specific_annot")
  
  # Remove rows where both values are 0
  df_sub <- diet_cell_types[! rowSums(diet_cell_types == 0) != 0, ]
  # print(df_sub)
  
  # Chisq test
  chisq <- chisq.test(as.matrix(df_sub))
  prop_tab <- prop.table(as.matrix(df_sub), 2)
  print(prop_tab)
  print(chisq)
  re_ls <- list("chisq_test" = chisq,
              "prop_tab" = prop_tab)
  return(re_ls)
})

names(chisq_ls) <- pairs_ls
```

Extract chisquare test from chisq_ls
```{r}
chisq_df <- purrr::map(chisq_ls, 1) %>%
  # Extract P value
  purrr::map(., "p.value") %>%
  data.frame(check.names = FALSE, row.names = "p-value") %>%
  t(.) %>%
  data.frame(check.names = FALSE) %>%
  tibble::rownames_to_column("pair") %>%
  tidyr::separate(col = pair,
                  into = c("experiment", "facs"),
                  sep = "-",
                  extra = "merge") %>%
  dplyr::arrange(facs, experiment) %>%
  dplyr::rename(p.value = `p-value`)

chisq_df$p.value_adj = p.adjust(p = chisq_df$p.value,
                                method = "bonferroni")
```

Visualize changes in all cell types across conditions
```{r fig.height=8, fig.width=13}
options(pillar.sigfig = 2)
# Add chisquare Pvalue to metadata
metadata %>%
  dplyr::left_join(chisq_df, by = c("experiment", "facs")) %>%
  dplyr::filter(experiment != "Normal") %>%
  # Only keep 1 pvalue
  dplyr::mutate(
    p.value_adj = signif(p.value_adj, digits = 3),
    p.value_txt = dplyr::if_else(specific_annot == "nerve-associated Fibroblasts",
                                         glue::glue("Adj p-value = {p.value_adj}"), NA_character_),
    specific_annot = stringr::str_wrap(specific_annot, 40)) %>%
  tibble::as_tibble() %>% 
  ggplot(data = .,
       aes(x = diet,
           y = count / total,
           color = specific_annot)) +
  geom_point(height = 0,
              width = 0.1) +
  geom_line(aes(group = specific_annot)) +
  facet_wrap(. ~ facs + experiment) +
  geom_text(aes(x = 1, y = 0.4, label = p.value_txt),
           size = 3,
           color = "#585858") +
  scale_y_log10() +
  theme_minimal() +
  labs(title = "Cell type proportion changes") +
  scale_color_manual(values = getPalette(colourCount)) +
  theme(legend.position = c(1, 0),
        legend.justification = c(1, 0))

```

Look visually if there are differences only in nerve-associated fibroblasts.
```{r fig.height=4, fig.width=5}
(
  ggplot(
     data = metadata %>% filter(specific_annot == 'nerve-associated Fibroblasts' &
                                  experiment != "Normal"),
     aes(
       x = diet,
       y = count / total,
       color = experiment,
       shape = facs))
  + geom_point(height = 0,
                width = 0.1)
  + geom_line(aes(group = pairs, linetype = facs))
  + scale_y_log10()
  + theme_minimal()
  + labs(subtitle = 'nerve-associated Fibroblasts')
)
```

Notes on how to get the OR, CI and P value can be found here: [How to obtain the P value from a confidence interval](https://doi.org/10.1136/bmj.d2304) and [Odds Ratio](https://www.ncbi.nlm.nih.gov/books/NBK431098/) \
Get Odds differences between conditions
```{r eval = FALSE, echo = FALSE, fig.height=10, fig.width=30}
OR_plots_ls <- lapply(pairs_ls, function(pair) {
  print(pair)
  # Extract sample id for those conditions
  sample_id_vec <- metadata %>%
    dplyr::filter(pairs == pair) %>%
    dplyr::pull(sample_id) %>%
    unique()
  print(sample_id_vec)
  
  # Subet both cell-type compositions
  diet_ctl <- metadata %>%
    dplyr::filter(sample_id == sample_id_vec[1]) %>%
    dplyr::mutate(odds_ctl_diet = count / other) %>%
    dplyr::rename(count_ctl = count,
                  other_ctl = other) %>%
    dplyr::select(specific_annot, odds_ctl_diet, count_ctl, other_ctl)
  
  diet_palm <- metadata %>%
    dplyr::filter(sample_id == sample_id_vec[2]) %>%
    dplyr::mutate(odds_palm_diet = count / other) %>%
    dplyr::rename(count_palm = count,
                  other_palm = other) %>%
    dplyr::select(specific_annot, odds_palm_diet, count_palm, other_palm)
  
  # join cell type numbers
  OR_cell_types <- diet_ctl %>%
    dplyr::left_join(diet_palm,
                     by = "specific_annot") %>%
    # remove cell type that is 0 in either conditions
    dplyr::filter_at(., .vars = c("count_palm"), dplyr::any_vars(. != 0)) %>%
    dplyr::filter_at(., .vars = c("count_ctl"), dplyr::any_vars(. != 0)) %>%
    dplyr::mutate(
      OR = odds_palm_diet / odds_ctl_diet,
      CI_low = exp(log(OR) - 1.96 * sqrt(1 / count_ctl + 1 / count_palm + 
                                         1 / other_ctl + 1 / other_palm)),
      CI_high = exp(log(OR) + 1.96 * sqrt(1 / count_ctl + 1 / count_palm + 
                                         1 / other_ctl + 1 / other_palm)),
      SE = (log(CI_high) - log(CI_low)) / (2 * 1.96),
      Z = log(OR) / SE,
      pval =  exp(-0.717 * abs(Z) - 0.416 * (abs(Z) ^ 2)),
      pval = signif(pval, digits = 2),
      pval_txt = glue::glue("P-value = {pval}"),
      specific_annot = stringr::str_wrap(string = specific_annot, width = 40)
      )
  # print(OR_cell_types)
  ggplot(data = OR_cell_types,
         aes(x = OR,
             y = specific_annot)) +
    geom_point() +
    geom_segment(aes(x = CI_low,
                     xend = CI_high,
                     yend = specific_annot)) +
    scale_x_log10() +
    geom_vline(xintercept = 1, color = "black", linetype = "dashed") +
    ggrepel::geom_text_repel(aes(label = pval_txt)) +
    theme_minimal() +
    labs(x = "Palm / Control",
         title = glue::glue("{pair} - Cell type OR change"))

})

cowplot::plot_grid(plotlist = OR_plots_ls,
                   align = "hv",
                   axis = "trbl")
```

Running Binomial Logistic Regression
```{r fig.height=10, fig.width=30}
OR_plots_ls <- lapply(pairs_ls, function(pair) {
  print(pair)
  # Extract sample id for those conditions
  sample_id_vec <- metadata %>%
    dplyr::filter(pairs == pair) %>%
    dplyr::pull(sample_id) %>%
    unique()
  print(sample_id_vec)
  
  # Detect which cell types have 0 counts and remove them...
  rm_ct <- metadata %>%
    dplyr::filter(sample_id %in% sample_id_vec & count == 0) %>%
    dplyr::pull(specific_annot) %>%
    unique
  
  # Remove those cell types since the logistic regression can't handle it
  df_a <- metadata %>%
    dplyr::filter(sample_id %in% sample_id_vec &
                    (! specific_annot %in% rm_ct))
  
  # Set formula 
  formula <- cbind(count, other) ~ specific_annot * sample_id
  model1 <- glm(formula = formula,
                family = 'binomial',
                data = df_a)
  
  emm1 <- emmeans::emmeans(model1,
                  specs = revpairwise ~ sample_id | specific_annot)
  c_results <- emm1$contrasts %>%
    summary(infer = TRUE, type = 'response') %>%
    rbind() %>%
    as.data.frame() %>%
    mutate(
      p.value = signif(p.value, digits = 2),
      pval_txt = glue::glue("P-value = {p.value}"),
      specific_annot = stringr::str_wrap(string = specific_annot, width = 40)
    )
    
  print(c_results)
  # Plot OR
  ggplot(data = c_results,
         aes(x = odds.ratio,
             y = specific_annot)) +
    geom_point() +
    geom_segment(aes(x = asymp.LCL,
                     xend = asymp.UCL,
                     yend = specific_annot)) +
    scale_x_log10() +
    geom_vline(xintercept = 1, color = "black", linetype = "dashed") +
    ggrepel::geom_text_repel(aes(label = pval_txt)) +
    theme_minimal() +
    labs(x = "Palm / Control",
         y = "",
         title = glue::glue("{pair} - Cell type OR change"))
})

OR_grid <- cowplot::plot_grid(plotlist = OR_plots_ls,
                   align = "hv",
                   axis = "trbl")

OR_grid

"{palm_28}/{plt_dir}/OR_epid20.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = OR_grid,
    base_height = 8,
    base_width = 12)
```

### DE genes
Next we are going to check DE genes between the same cell type across conditions - 1st xenograph vs 2nd xenograph
```{r eval = FALSE}
DE_df <- lapply(c("PLKO", "shCD36", "shGal", "Plko", "_shcD36_"), function(cond) {
  print(cond)
  keep <- stringr::str_detect(string = annot_se@meta.data[, "sample_id"],
                              pattern = cond)
  tmp_se <- annot_se[, keep]
  
  Seurat::Idents(tmp_se) <- tmp_se@meta.data[, "sample_id"]
  diets <- sort(unique(tmp_se$sample_id))
  print(diets)
  
  ct_ls <- unique(tmp_se$specific_annot)
  df <- lapply(ct_ls, function(ct) {
    print(ct)
    se_sub <- tmp_se[, tmp_se$specific_annot == ct]
    
    if (ncol(table(se_sub@meta.data[, "specific_annot"],
                   se_sub@meta.data[, "sample_id"])) == 2) {
      markers <- Seurat::FindMarkers(object = se_sub,
                                   ident.1 = diets[1],
                                   ident.2 = diets[2])
    } else {
      markers <- data.frame(p_val = 1,
                            avg_logFC = 0,
                            pct.1 = 0,
                            pct.2 = 0,
                            p_val_adj = 1)
    }
    
    markers$cell_type <- ct
    markers$comparison <- sprintf("%s vs %s", diets[1], diets[2])
    markers <- markers %>%
      tibble::rownames_to_column("gene")
    
    return(markers)
    # de_ls[[cond]][[ct]] <- markers
  }) %>%
    dplyr::bind_rows()
}) %>% 
  dplyr::bind_rows()

saveRDS(object = DE_df,
        file = here::here(sprintf("%s/%s/DE_diet.rds",
                                  anot_28, robj_dir)))
```

```{r}
DE_df <- readRDS(file = here::here(sprintf("%s/%s/DE_diet.rds",
                                  anot_28, robj_dir)))

DT::datatable(data = DE_df,
              filter = "top")
```

Check if there are DE expressed genes in all cell types of the same condition
```{r}
intersect_genes <- lapply(unique(DE_df$comparison), function(cond) {
  lapply(unique(DE_df$cell_type), function(ct){
    DE_df %>%
      dplyr::filter(comparison == cond & cell_type == ct) %>%
      dplyr::pull("gene")
  }) %>% 
      purrr::reduce(., intersect)
  
})

names(intersect_genes) <- unique(DE_df$comparison)
```

Show intersecting DE genes between the cell types of the same condition.
```{r}
intersect_genes_mod <- purrr::map(intersect_genes, ~ glue::glue_collapse(., ", "))
max_len <- max(lengths(intersect_genes_mod))
df <- do.call(cbind.data.frame, c(lapply(intersect_genes_mod, function(x) 
               c(x, rep('', max_len - length(x)))), stringsAsFactors = FALSE))
knitr::kable(df)
```

And now check if there are genes DE expressed intersecting across all conditions
```{r}
purrr::reduce(intersect_genes, intersect)
```

We also check if between the same comparison but in 2 different replicates we observe the same DE genes
PLKO vs Plko
```{r fig.height=6, fig.width=12, eval = FALSE}
intersect(intersect_genes[["25-PLKOControlDiet vs 25-PLKOPalmDiet"]],
          intersect_genes[["25_Plko_ct_diet vs 25_Plko_Palmdiet"]])

ggVennDiagram(x = intersect_genes[c("25-PLKOControlDiet vs 25-PLKOPalmDiet",
                                    "25_Plko_ct_diet vs 25_Plko_Palmdiet")])
```

shcD36 vs \_shcD36_
```{r fig.height=6, fig.width=12, eval = FALSE}
intersect(intersect_genes[["25-shCD36ControlDiet vs 25-shCD36PalmDiet"]],
          intersect_genes[["25_shcD36_ctdiet vs 25_shcD36_PALM_diet"]])

ggVennDiagram(x = intersect_genes[c("25-shCD36ControlDiet vs 25-shCD36PalmDiet",
                                    "25_shcD36_ctdiet vs 25_shcD36_PALM_diet")])
```

Lastly we want to observe the intersection of DE genes for cell types of interest between replicates so as to determine which genes are DE expressed in both conditions.
```{r}
# Define comparisons of interest
comparisons <- list()
comparisons[["PLKO"]] <- c("25-PLKOControlDiet vs 25-PLKOPalmDiet",
                           "25_Plko_ct_diet vs 25_Plko_Palmdiet")

comparisons[["shCD36"]] <- c("25-shCD36ControlDiet vs 25-shCD36PalmDiet",
                             "25_shcD36_ctdiet vs 25_shcD36_PALM_diet")

# Define cell types of interest
ct_interest <- c("nerve-associated Fibroblasts",
                 "Retinol uptaking basal cells",
                 "Lymphoid Endothelial Cells")
```

Show the intersecting DE genes in a table
```{r}
intersect_int <- lapply(comparisons, function(cond_ls) {
  tmp_df <- lapply(ct_interest, function(ct){
    repl1 <- DE_df %>%
      dplyr::filter(comparison == cond_ls[[1]] & cell_type == ct) %>%
      dplyr::pull("gene")
    
    repl2 <-  DE_df %>%
      dplyr::filter(comparison == cond_ls[[2]] & cell_type == ct) %>%
      dplyr::pull("gene")
    
    inter <- dplyr::intersect(repl1, repl2)
    glue::glue_collapse(inter, ", ")
  })
  names(tmp_df) <- ct_interest
  return(tmp_df)
})

max_len <- max(lengths(intersect_int))
df_int <- do.call(cbind.data.frame, c(lapply(intersect_int, function(x) 
               c(x, rep("", max_len - length(x)))), stringsAsFactors = FALSE))

knitr::kable(df_int)
```

Show the results in a VennDiagram
```{r fig.height=7, fig.width=21, eval = FALSE}
intersect_venn <- lapply(comparisons, function(cond_ls) {
  tmp_venn <- lapply(ct_interest, function(ct){
    repl1 <- DE_df %>%
      dplyr::filter(comparison == cond_ls[[1]] & cell_type == ct) %>%
      dplyr::pull("gene")
    
    repl2 <-  DE_df %>%
      dplyr::filter(comparison == cond_ls[[2]] & cell_type == ct) %>%
      dplyr::pull("gene")
    
    tmp_ls <- list(repl1, repl2)
    names(tmp_ls) <- cond_ls
    ggVennDiagram(x = tmp_ls)
  })
})

cowplot::plot_grid(plotlist = intersect_venn[[1]], nrow = 1)
cowplot::plot_grid(plotlist = intersect_venn[[2]], nrow = 1)
```

## Session Info
```{r}
sessionInfo()
```
