---
title: "07-Join Annotation"
author: "Marc Elosua"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  BiocStyle::html_document:
    toc: true
    toc_float: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, out.width = "100%", fig.align='center',
                      message = FALSE, warning = FALSE, fig.width = 8)
# options(width = 1200)
```

## Introduction
In this Rmarkdown document we are going to join the annotations and plot marker gene violin plots.

## Libraries
```{r warning = FALSE, message = FALSE}
library(Seurat)
library(ggpubr)
library(cowplot)
library(dplyr)
library(ggplot2)
library(stringr)
library(readr)
library(plotly)
```

## Parameters
Here we will define and load objects and functions that will be used throughout the document.
```{r}
source(here::here("misc/paths.R"))
```

## Load data
The data used in this Rmarkdown document comes from 03-integration_clustering/04-EPID_integration.Rmd
```{r}
# General annotation
merged_se <- "{anot_28}/{robj_dir}/harmony_se_general.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)

# Specific annotation
sp_annot <- "{anot_28}/{robj_dir}/specific_annotation.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)
table(sp_annot$specific_annot)
```

## Analysis
Join the specific annotation to the seurat object
```{r}
merged_se@meta.data <- merged_se@meta.data %>%
  tibble::rownames_to_column("barcode") %>%
  dplyr::left_join(sp_annot %>%
                     tibble::rownames_to_column("barcode"),
                   by = "barcode") %>%
  dplyr::mutate(
    specific_annot = dplyr::if_else(is.na(specific_annot),
                                    general_annot,
                                    specific_annot)
    # specific_annot = general_annot
  ) %>%
  tibble::column_to_rownames("barcode")
table(merged_se$specific_annot)
table(merged_se$general_annot)

```

Add a bit of cluster substructure in the basal cluster for the deconvolution:
```{r}
Seurat::DimPlot(object = merged_se[, merged_se$specific_annot == "Basal Cells"], group.by = "SCT_snn_res.0.25")
Seurat::DimPlot(object = merged_se, group.by = c("SCT_snn_res.0.25", "specific_annot"), label = TRUE)

merged_se@meta.data <- merged_se@meta.data %>%
  dplyr::mutate(
    specific_annot_mod = dplyr::case_when(
      SCT_snn_res.0.25 == 0 ~ "Basal Cells - 1",
      SCT_snn_res.0.25 == 1 ~ "Proliferating Basal Cells",
      SCT_snn_res.0.25 == 2 ~ "Basal Cells - 2",
      SCT_snn_res.0.25 == 3 ~ "Basal Cells - 3",
      SCT_snn_res.0.25 == 4 ~ "Lymphoid Endothelial",
      SCT_snn_res.0.25 == 5 ~ "Erithrocytes",
      SCT_snn_res.0.25 == 6 ~ "Vascular Endothelial",
      SCT_snn_res.0.25 == 7 ~ "Basal Cells - 4",
      SCT_snn_res.0.5 == 14 ~ "Tumour-associated Schwann Cells",
      SCT_snn_res.0.5 == 15 ~ "SMC", 
      SCT_snn_res.0.25 == 9 ~ "Basal Cells - 5",
      SCT_snn_res.0.25 == 10 ~ "Erithrocytes",
      )
    )
Seurat::DimPlot(object = merged_se, group.by = c("specific_annot_mod"), label = TRUE)

```

Since ribosomal are poor quality cells we remove them
```{r}
merged_se <- subset(merged_se, subset = specific_annot != "Ribosomal")
```

Look at the new annotation
```{r}
Seurat::DimPlot(
  merged_se,
  group.by = "specific_annot")
```

### Violin plots
Next we are going to plot Violin plots for the expression of marker genes that differentiate the cell types in this dataset
```{r}
ct_order <- c("Basal Cells", "Proliferating Basal Cells", "Lymphoid Endothelial",
              "Vascular Endothelial", "Tumour-associated Schwann Cells", "SMC",
              "Erithrocytes")

gene_ls <- c("Krt14", "Krt5", # Basal epidermis cells
             "Top2a", "Mki67", # Proliferation
             "Ccl21a", "Lyve1", # Lymphoid endothelial
             "Flt1", "Vcam1", # Vascular endothelial
             "Apod", "Ecm1", # Tumour-associated Schwann cells
             "Tagln", "Acta2", # SMC
             "Hba-a1", "Hbb-bt") # Erithrocytes

count_int <- Seurat::GetAssayData(
  object = merged_se,
  assay = "SCT",
  slot = "data") %>%
  .[gene_ls, ] %>%
  as.matrix(.) %>%
  t() %>%
  data.frame(.,check.names = FALSE) %>% 
  tibble::rownames_to_column("barcode")

vln_plt <- merged_se@meta.data %>%
  dplyr::select(specific_annot) %>%
  tibble::rownames_to_column("barcode") %>%
  dplyr::left_join(count_int, by = "barcode") %>%
  tibble::column_to_rownames("barcode") %>%
  tidyr::pivot_longer(
    cols = dplyr::all_of(gene_ls),
    names_to = "gene",
    values_to = "expr") %>%
  dplyr::mutate(
    gene = factor(gene, levels = gene_ls),
    specific_annot = factor(specific_annot, levels = ct_order)
  ) %>%
  ggplot2::ggplot(., 
                  aes(
                    x = specific_annot,
                    y = expr,
                    fill = specific_annot,
                    color = specific_annot)) +
  ggplot2::geom_violin() +
  ggplot2::facet_wrap(. ~ gene, ncol = 2, scales = "free_y") +
  ggplot2::labs(fill = "", color = "", y = "Normalized Expression", x = "") +
  ggplot2::theme_classic() +
  ggplot2::theme(
     strip.background = ggplot2::element_blank(),
     strip.text = ggplot2::element_text(face = "bold"),
     axis.text.x = ggplot2::element_text(
       angle = 60,
       hjust = 1,
       vjust = 1))

vln_plt

"{anot_28}/{plt_dir}/vln_plot_markers_epid20.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = vln_plt,
    base_height = 20,
    base_width = 12)
```

### Feature plots
We will also plot the feature plots of these genes
```{r}
library(RColorBrewer)
feat_plt <- Seurat::FeaturePlot(object = merged_se,
                    features = gene_ls) &
  Seurat::NoLegend() & 
  ggplot2::scale_colour_gradientn(
    colours = RColorBrewer::brewer.pal(n = 11, name = "YlOrRd"))

"{anot_28}/{plt_dir}/feat_plot_markers_epid20.pdf" %>%
  glue::glue() %>%
  here::here() %>%
  cowplot::save_plot(
    filename = .,
    plot = feat_plt,
    base_height = 15,
    base_width = 15)

```

### Save marker genes
First of all we need to compute the markers for the cell types using `Seurat's` funtion `FindAllMarkers`.
```{r}
Seurat::Idents(merged_se) <- merged_se$specific_annot
sc_markers <- Seurat::FindAllMarkers(
  object = merged_se,
  assay = "SCT",
  slot = "data",
  only.pos = TRUE)

"{anot_28}/{robj_dir}/markes_specific_epid20.rds" %>%
  glue::glue() %>%
  here::here() %>%
  saveRDS(
    object = sc_markers,
    file = .)
```

Load the markers so we don't have to compure them redundantly
```{r eval = FALSE}
sc_markers <- "{anot_28}/{robj_dir}/markes_specific_epid20.rds" %>%
  glue::glue() %>%
  here::here() %>%
  readRDS(file = .)

# sc_markers <- readRDS(file = here::here(glue::glue("{map_27}/{robj_dir}/
#                                             markes_specific.rds")))

markers_fn <- "{anot_28}/{robj_dir}/markers_update.xlsx" %>%
  glue::glue() %>%
  here::here()

# Write results to spreadsheet, each cluster in a new sheet
lapply(unique(sc_markers$cluster), function(clust) {
  xlsx::write.xlsx(
    x = sc_markers %>%
      dplyr::filter(cluster == clust) %>%
      dplyr::arrange(dplyr::desc(avg_log2FC)),
    file = markers_fn,
    sheetName = glue::glue("Cluster {clust}"),
    row.names = FALSE,
    append = file.exists(markers_fn))
})
```

## Save data
```{r}
"{anot_28}/{robj_dir}/harmony_se_annot.rds" %>%
  glue::glue() %>%
  here::here() %>%
  saveRDS(
    object = merged_se,
    file = .)

```

## Session Info
```{r}
sessionInfo()
```